% Blueprint content — includes all chapter files.
% This file is used by both web.tex and print.tex.

\chapter{Formal Power Series}
\input{chapter_notations}
\input{chapter_commutative_rings}
\input{chapter_fps_definition}
\input{chapter_dividing_fps}
\input{chapter_polynomials}
\input{chapter_substitution}
\input{chapter_derivatives}
\input{chapter_exp_log}
\input{chapter_non_integer_powers}
\input{chapter_integer_compositions}
\input{chapter_xn_equivalence}
\input{chapter_infinite_products_1}
\input{chapter_infinite_products_2}
\input{chapter_weighted_sets}
\input{chapter_limits}
\input{chapter_laurent_series}
\input{chapter_multivariate}

\chapter{Integer Partitions}
\input{chapter_partition_basics}
\input{chapter_pentagonal_jacobi}
\input{chapter_qbinomial_basic}
\input{chapter_qbinomial_formulas}

\chapter{Permutations}
\input{chapter_permutation_basics}
\input{chapter_inversions_1}
\input{chapter_inversions_2}
\input{chapter_signs}
\input{chapter_cycle_decomposition}

\chapter{Signed Counting}
\input{chapter_alternating_sums}
\input{chapter_inclusion_exclusion_1}
\input{chapter_boolean_mobius}
\input{chapter_subtractive_methods}

\chapter{Determinants}
\input{chapter_determinants_basic}
\input{chapter_cauchy_binet}
\input{chapter_desnanot_jacobi}
\input{chapter_lgv_1}
\input{chapter_lgv_2}

\chapter{Symmetric Functions}
\input{chapter_symmetric_definitions}
\input{chapter_monomial_symmetric}
\input{chapter_schur_basics}
\input{chapter_littlewood_richardson}
\input{chapter_pieri_jacobi_trudi}

\appendix
\chapter{Details}
\input{chapter_details_infprod_1}
\input{chapter_details_infprod_2}
\input{chapter_details_domino}
\input{chapter_details_limits}
\input{chapter_details_laurent}

\begin{thebibliography}{999999999}

\bibitem[19s]{19s}Darij Grinberg, \textit{Introduction to Modern Algebra (UMN
Spring 2019 Math 4281 notes)}, 29 June 2019.\newline\url{http://www.cip.ifi.lmu.de/~grinberg/t/19s/notes.pdf}

\bibitem[19s-mt3s]{19s-mt3s}Darij Grinberg, \textit{Math 4281: Introduction to
Modern Algebra, Spring 2019: Midterm 3 with solutions}.\newline\url{https://www.cip.ifi.lmu.de/~grinberg/t/19s/mt3s.pdf}

\bibitem[19fco]{19fco}Darij Grinberg, \textit{Enumerative Combinatorics: class
notes (Drexel Fall 2019 Math 222 notes)}, 11 September 2022.\newline\url{http://www.cip.ifi.lmu.de/~grinberg/t/19fco/n/n.pdf}

\bibitem[Ford21]{Ford21}Timothy J. Ford, \textit{Abstract Algebra}, draft of a
book, 14 August 2024.\newline\url{https://tim4datfau.github.io/Timothy-Ford-at-FAU/preprints/Algebra_Book.pdf}

\bibitem[Grinbe15]{detnotes}Darij Grinberg, \textit{Notes on the combinatorial
fundamentals of algebra}, 15 September 2022.\newline
\url{http://www.cip.ifi.lmu.de/~grinberg/primes2015/sols.pdf}

\bibitem[Grinbe17]{logexp}Darij Grinberg, \textit{Why the log and exp series
are mutually inverse}, 11 May 2018.\newline\url{https://www.cip.ifi.lmu.de/~grinberg/t/17f/logexp.pdf}

\bibitem[Laue15]{Laue-det}Hartmut Laue, \textit{Determinants}, version 17 May
2015,\newline
\url{http://www.math.uni-kiel.de/algebra/laue/homepagetexte/det.pdf}

\bibitem[Strick13]{Strick13}Neil Strickland, \textit{MAS201 Linear Mathematics
for Applications}, lecture notes, 11 February 2020.\newline
\url{https://neilstrickland.github.io/linear_maths/notes/linear_maths.pdf}

\bibitem[Zeilbe85]{Zeilbe}
\href{http://www.math.rutgers.edu/~zeilberg/mamarimY/DM85.pdf}{Doron
Zeilberger, \textit{A combinatorial approach to matrix algebra}, Discrete
Mathematics 56 (1985), pp. 61--72.}

\bibitem[20f]{20f}Darij Grinberg, \textit{Math 235: Mathematical Problem
Solving}, 22 March 2021.\newline\url{http://www.cip.ifi.lmu.de/~grinberg/t/20f/mps.pdf}

\bibitem[Benede25]{Benede25}Bruno Benedetti, \textit{Introduction to Abstract
Algebra \textquotedblleft Rings First\textquotedblright}, lecture notes, 6
March 2025.\newline\url{https://www.math.miami.edu/~bruno/algebra2.pdf}

\bibitem[BluCos16]{BluCos16}\href{https://arxiv.org/abs/1301.7116v5}{Ben
Blum-Smith, Samuel Coskey, \textit{The Fundamental Theorem on Symmetric
Polynomials: History's First Whiff of Galois Theory}, arXiv:1301.7116v5.}

\bibitem[Bresso99]{Bresso99}%
\href{https://doi.org/10.1017/CBO9780511613449}{David M. Bressoud,
\textit{Proofs and Confirmations: The Story of the Alternating Sign Matrix
Conjecture}, Cambridge University Press 1999}.\newline See
\url{http://web.archive.org/web/20220531023303/https://www.macalester.edu/~bressoud/books/PnC/PnCcorrect.html}

\bibitem[Dumas08]{Dumas08}Fran\c{c}ois Dumas, \textit{An introduction to
noncommutative polynomial invariants}, lecture notes (Cimpa-Unesco-Argentina
\textquotedblleft Homological methods and representations of non-commutative
algebras\textquotedblright, Mar del Plata, Argentina March 6 - 17,
2006).\newline\url{https://lmbp.uca.fr/~fdumas/fichiers/CIMPA.pdf}

\bibitem[Goodma15]{Goodman}Frederick M. Goodman, \textit{Algebra: Abstract and
Concrete}, edition 2.6, 1 May 2015.\newline%
\url{http://homepage.math.uiowa.edu/~goodman/algebrabook.dir/book.2.6.pdf} .

\bibitem[Loehr11]{Loehr-BC}%
\href{http://www.math.vt.edu/people/nloehr/bijbook.html}{Nicholas A. Loehr,
\textit{Bijective Combinatorics}, Chapman \& Hall/CRC 2011.}

\bibitem[Smith95]{Smith95}\href{https://doi.org/10.1201/9781439864470}{Larry
Smith, \textit{Polynomial Invariants of Finite Groups}, A K Peters 1995.}

\bibitem[Bourba74]{Bourba74}%
\href{http://libgen.rs/book/index.php?md5=3270565F6D0052635A1550883588204C}{Nicolas
Bourbaki, \textit{Algebra I: Chapters 1--3}, Addison-Wesley 1974}.

\bibitem[Zeilbe98]{zeilberger-twotime}Doron Zeilberger, \textit{Dodgson's
Determinant-Evaluation Rule proved by Two-Timing Men and Women}, The
Electronic Journal of Combinatorics, vol. 4, issue 2 (1997) (The Wilf
Festschrift).

\end{thebibliography}
